On the minimum leaf conductance: its role in models of plant water use, and ecological and environmental controls

### 2017-09-21 manuscript.Rmd
some more writing; table of contents Tansley style

```{r include=FALSE}
includefigs <- F
library(png)
library(grid)

@@

@@

\
The absolute minimum attainable water loss rate through the cuticle only, which is typically measured on isolated cuticles of the adaxial (non-stomatal) side of the leaf. Other measurements allow for the estimation of water loss through incompletely closed and broken stomata. Of particular interest is the method where leaves are detached from the plant, and their weight loss monitored over time. This method (mass loss of detached leaves, MLD) is interesting because it more closely simulates actual field conditions for plants that are severely droughted, and includes all surfaces of the leaf. Since the seminal review by @kerstiens1996, no attempt has been made to synthesize the literature for values of g~min~, which is timely as models start to incorporate more detailed drought dynamics all the way to the point of mortality by desiccation (Xu, Blackman, Martin-StPaul, Sperry?). Instead, many models rely on inaccurately estimated 'minimum' conductance from cuvette data, by applying regression techniques. We critically examine this method in Section III.

\
Previews reviews have focused on the biology of the plant cuticle, and detailed physiology and anatomy of water transport across cuticles (riederer, kerstiens). Thus there is a disconnect between dizzying level of detailed understanding of the plant cuticle [@riederer2008], and the very simple  assumptions made when applying ecosystem models. We aim to improve the connection somewhat, and point at obvious gaps in understanding.

@@

Measurements of (presumed) minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories. The absolute minimum conductance attained by leaves is that through the cuticle only. Measurements of cuticular conductance (g~cut~) are typically made on isolated cuticles [@riederer2001] (g~cut,isol~), or on the side of the leaf that lacks stomata ('adaxial' side), or by sealing the side with stomata (g~cut,seal~). Another measurement is to detach leaves, and measure leaf weight over time as the leaf dries out. After some time, a constant minimum transpiration rate is attained (sometimes described as the 'residual transpiration rate'), from which minimum conductance (g~min~) can be inferred using measured vapour pressure deficit [@sinclair1986]. This method most closely resembles conditions that plants may experience during a dry-down or periods of extreme stress.

\
For the @lin2015 database, we fit the model of @medlyn2011 using non-linear regression, while @miner2017 presented results from linear regression of the @ballberry1987 model. Measurements of conductance of isolated, non-stomatal cuticles (g~cut,isol~), and of conductance of detached leaves with the stomatal side sealed (g~cut,seal~) were taken from the compilation by @kerstiens1996.  Data on minimum conductance (g~min~) from weight loss of detached leaves was compiled from 44+ original papers, see SuppInfo A for a list of papers and additional figures describing this new compilation. For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources. All conductance values were converted to per unit projected surface area. Units were converted to molar fluxes assuming 41 mol m^-3^ (cf. @kerstiens1996). Although it is quite clear that night-time conductance (g~dark~) is much larger than a true 'minimum' conductance, it is frequently used as an independent measurement of g~0~ [@lombardozzi2017], and thus a comparison of magnitudes is useful. A compilation of night-time conductance (g~dark~) values was taken from @lombardozzi2017, who compiled cuvette (and sometimes sap flow-based) measurements of ca. 150 species during night-time. Actual minimum daytime values were estimated from the Lin et al. database, by taking those data where A~n~ < 2 $mu$ mol m^-2^ s^-1^, and calculating the minimum g~s~ for each of the XXX studies included in that dataset. This estimate was added as a reference as it represents an actual minimum values measured, rather than a fitted/extrapolated g~0~ term.

\
Conductance in the dark was on average eight-fold higher than g~min~. It is no new observation that stomata do not close in the dark, but do close on detached leaves, as already demonstrated by @hygen1954 in Scots pine, and described by @koerner1994. More recently, @walden-coleman2013 showed much higher g~dark~ compared to g~min~ in Cotton. However, @cavender-bares2007 gdark similar to gmin in oaks. Actual measured stomatal conductance during conditions when photosynthesis was low (g~s,low A~) was also much lower than fitted g~0~ or g~dark~. In other words, actual daytime conductance is much lower than the presumed minimum estimated from fitted g~0~. A similar result was already presented by @dekauwe2015 (their  Fig. A2).

# III. Estimating minimum conductance from cuvette data

The standard approach in estimating g~0~ for use in Eq. 1 is to use regression with g~s~ as the response variable, and the right-hand side of Eq. 1 as the predictor. The g~0~ parameter is thus estimated as the intercept. Here we briefly discuss some statistical aspects of this estimation procedure, and draw the general conclusion that g~0~ is poorly estimated in this way. The difficulty of the estimation can already be concluded from the fact that a) many studies set g~0~ to some assumed value rather than fitting it (e.g. @leuning1995 uses 0.01 mol m^-2^ s^-1^ for all species); presumably because unreliable estimates were obtained, b) negative g~0~ estimates from regression are commonly reported [@leuning1995; @heroult2013; @miner2017], though nonsensical.

@@

Discuss Martin-StPaul (2017); key role of minimum conductance in a model of desiccation dynamics

# VI. Cuticular conductance

@@

\
It is important to bear in mind other functions of the cuticle, including attenuation of radiation, as a barrier for fungal pathogens, and various interactions with insects [@mueller2008] [@kerstiens1996c;@riederer2008]. The thickness of the cuticle per se may also have to do with mechanical strength, as @onoda2012 reported that thicker cuticles are more resistant to tearing (the force to tear and tensile strength both approx proportional to cuticle thickness). Thick cuticles may also be useful because  exposure to wind has been shown to dislodge cuticular compounds by absrasion [@rogge1993], thereby dramatically increasing the conductance of the cuticle [@hadley1983;@grace1974;@vangardingen1991]. Similarly, exposure to simulated rain damages the cuticle and increases its conductance [@baker1986]. Such exposure to the elements may be responsible for the typical gradual increase in g~min~ with leaf age in field-grown plants [@jordan2007].

@@

# VII. Nighttime conductance

@@

\
First, ample observations suggect that g~dark~ is not a fixed rate but varies tremendously during the night. @caird2007 (and references therein) described how for many species, g~dark~ is not stable throughout the night period. Instead, endogenous, gradual increases in stomatal opening during predawn hours have been reported in many species under natural field conditions as well as in controlled environments [@rawson1988]. @rescodedios2016 showed that, in *Eucalyptus camaldulensis*, g~dark~ in the period just after sunset was much lower than pre-dawn g~dark~. Strong evidence for endogenous regulation of g~dark~ was reported by @rescodedios2013, who showed that g~dark~ fluctuated throughout the night despite environmental conditions being held constant in whole-tree chambers. Clearly g~dark~ is an actively controlled process that cannot be adequately summarized by a single constant g~0~.

# VIII. Environmental and ecological variation in minimum conductance

@@

# Conclusions

>>>>

```{r include=FALSE}
includefigs <- F
library(png)
library(grid)
library(knitr)

@@

**Contents:**

```{r, echo=FALSE}
tab <- read.csv(text="I. Introduction
II. Comparison of various definitions of minimum conductance
III. Problems with estimating minimum conductance from cuvette data
IV. Use of minimum conductance in models
\t1. Models of water-use efficiency
\t2. Models of drought response of stomatal conductance
V. Cuticular conductance
VI. Nighttime conductance
VII. Environmental and ecological variation in minimum conductance 
\t1. Contribution of stomata to minimum conductance
\t2. Minimum conductance shows little meaningful interspecific variation
\t3. Acclimation to drought
\t4. Acclimation and instantaneous response to temperature
\t5. Other factors
VIII. Conclusions", header=F)

knitr::kable(tab[[1]], col.names="")

```

@@

\
The absolute minimum attainable water loss rate through the cuticle only, which is typically measured on isolated cuticles of the adaxial (non-stomatal) side of the leaf. Other measurements allow for the estimation of water loss through incompletely closed and broken stomata. Of particular interest is the method where leaves are detached from the plant, and their weight loss monitored over time. This method (mass loss of detached leaves, MLD) is interesting because it more closely simulates actual field conditions for plants that are severely droughted, and includes all surfaces of the leaf. Since the seminal review by @kerstiens1996, no attempt has been made to synthesize the literature for values of g~min~, which is timely as models start to incorporate more detailed drought dynamics all the way to the point of mortality by desiccation (Xu, Blackman, Martin-StPaul, Sperry?). Instead, many models rely on inaccurately estimated 'minimum' conductance from cuvette data, by applying regression techniques.

\
(*expand*) Previews reviews have focused on the biology of the plant cuticle, and detailed physiology and anatomy of water transport across cuticles (riederer, kerstiens). Thus there is a disconnect between dizzying level of detailed understanding of the plant cuticle [@riederer2008], and the very simple  assumptions made when applying ecosystem models. We aim to improve the connection somewhat, and point at obvious gaps in understanding. Throughout our review, we discuss both the practical implementation of a 'minimum' conductance g~0~ as implemented in e.g. Ball-Berry models of stomatal conductance (Eq. 1) and similar, and a 'true' minimum conductance as estimated on detached leaves (g~min~).

@@

Measurements of (presumed) minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories. The absolute minimum conductance attained by leaves is that through the cuticle only. Measurements of cuticular conductance (g~cut~) are typically made on isolated non-stomatal cuticles [@riederer2001] (g~cut,isol~), or via gas exchange techniques, or mass loss of detached leaves vby sealing the side with stomata (g~cut,seal~). The method we here refer to as minimum conductance is that of mass loss of detached leaves (MLD). In this method, leaves are detached and leaf mass is monitored over time as the leaf dries out. After some time, a constant minimum transpiration rate is attained (sometimes described as the 'residual transpiration', or 'epidermal transpiration'), from which minimum conductance (g~min~) can be inferred using measured vapour pressure deficit [@sinclair1986]. This method most closely resembles conditions that plants may experience during a dry-down or periods of extreme stress. The g~0~ parameter, which is a *modelled* minimum conductance reached when photosynthesis rate is zero, is typically estimated by applying regression to Eq. 1 (of which various forms exist).

\ 
We quantitatively compared each of these different 'minimum' codnductance values by synthesizing existing data. To estimate g~0~ from regression, we used the @lin2015 database to fit the model of @medlyn2011 using non-linear regression, and data presented by @miner2017 on the linear regression of the @ballberry1987 model. Measurements of conductance of isolated, non-stomatal cuticles (g~cut,isol~), and of conductance of detached leaves with the stomatal side sealed (g~cut,seal~) were taken from the compilation by @kerstiens1996.  Data on minimum conductance (g~min~) from MLD was compiled from 44+ original papers (see SuppInfo A for a list of papers and additional figures describing this new compilation). For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources. All conductance values were converted to per unit projected surface area. Units were converted to molar fluxes assuming 41 mol m^-3^ (cf. @kerstiens1996). Although it is quite clear that night-time conductance (g~dark~) is much larger than a true 'minimum' conductance, it is frequently used as an independent measurement of g~0~ [@lombardozzi2017], and thus a comparison of magnitudes is useful. A compilation of night-time conductance (g~dark~) values was taken from @lombardozzi2017, who compiled cuvette (and sometimes sap flow-based) measurements of ca. 150 species during night-time. Actual minimum daytime values were estimated from the Lin et al. database, by taking those data where A~n~ < 2 $mu$ mol m^-2^ s^-1^, and calculating the minimum g~s~ for each of the XXX studies included in that dataset. This estimate was added as a reference as it represents an actual minimum values measured, rather than a fitted/extrapolated g~0~ term.

\
(*expand*) The results are presented in Fig. 1. Conductance in the dark was on average eight-fold higher than g~min~. It is no new observation that stomata do not close in the dark, but do close on detached leaves, as already demonstrated by @hygen1954 in Scots pine, and described by @koerner1994. More recently, @walden-coleman2013 showed much higher g~dark~ compared to g~min~ in Cotton. However, @cavender-bares2007 gdark similar to gmin in oaks. Actual measured stomatal conductance during conditions when photosynthesis was low (g~s,low A~) was also much lower than fitted g~0~ or g~dark~. In other words, actual daytime conductance is much lower than the presumed minimum estimated from fitted g~0~. A similar result was already presented by @dekauwe2015 (their  Fig. A2).

# III. Problems with estimating minimum conductance from cuvette data

The standard approach in estimating g~0~ for use in Eq. 1 is to use regression with g~s~ as the response variable, and the right-hand side of Eq. 1 as the predictor. The g~0~ parameter is thus estimated as the intercept. Here we briefly discuss some statistical aspects of this estimation procedure, and draw the general conclusion that g~0~ is poorly estimated by this method. The difficulty of the estimation can already be concluded from the fact that a) many studies set g~0~ to some assumed value rather than fitting it (e.g. @leuning1995 uses 0.01 mol m^-2^ s^-1^ for all species); presumably because unreliable estimates were obtained, b) negative g~0~ estimates from regression are commonly reported [@leuning1995; @heroult2013; @miner2017], though it is clearly non-sensical to suggest negative conductance values.

@@

Discuss Martin-StPaul (2017); key role of minimum conductance in a model of desiccation dynamics.

**Could we have a figure showing time to mortality as a function of gmin? I know Nico wants to use this in a separate paper, but it might draw some attention here.**

# V. Cuticular conductance

@@

\
It is important to bear in mind other functions of the cuticle, including attenuation of radiation, as a barrier for fungal pathogens, and various interactions with insects [@mueller2008] [@kerstiens1996c;@riederer2008]. The thickness of the cuticle per se may also have to do with mechanical strength, as @onoda2012 reported that thicker cuticles are more resistant to tearing (the force to tear and tensile strength both approx proportional to cuticle thickness). Thick cuticles may also be useful because  exposure to wind has been shown to dislodge cuticular compounds by absrasion [@rogge1993], thereby dramatically increasing the conductance of the cuticle [@hadley1983;@grace1974;@vangardingen1991]. Similarly, exposure to simulated rain damages the cuticle and increases its conductance [@baker1986], most dramatically demonstrated in an ice storm, leading to much higher g~min~ [@boyce2003]. Such exposure to the elements may be responsible for the typical gradual increase in g~min~ with leaf age in field-grown plants [@jordan2007].

@@

# VI. Nighttime conductance

@@

\
First, ample observations suggect that g~dark~ is not a fixed rate but varies tremendously during the night. @caird2007 (and references therein) described how for many species, g~dark~ is not stable throughout the night period. Instead, endogenous, gradual increases in stomatal opening during predawn hours have been reported in many species under natural field conditions as well as in controlled environments [@rawson1988]. @rescodedios2016 showed that, in *Eucalyptus camaldulensis*, g~dark~ in the period just after sunset was much lower than pre-dawn g~dark~. Strong evidence for endogenous regulation of g~dark~ was reported by @rescodedios2013, who showed that g~dark~ fluctuated throughout the night despite environmental conditions being held constant in whole-tree chambers. In addition, g~dark~ can show a clear response to vapour pressure deficit [@barbour2007] during night-time conditions. Clearly g~dark~ is an actively controlled process that cannot be adequately summarized by a single constant g~0~.

# VII. Environmental and ecological variation in minimum conductance

@@

# VIII. Conclusions
